---
title: "GTEx rank issue and imputation problem"
author: "Kushal K. Dey and Rahul Mazumder"
date: "11/25/2019"
output: html_document
---

```{r}
library(ggplot2)
library(corrplot)
library(softImpute)
```

In this script, we show that the GTEx tissue-wide gene expression data across individuals does not have a low rank factorization by looking at the eigen-values of the pairwise sample correlation matrix for a gene (PLIN1) from the GTEx genes.

We also show that imputation based models such as SoftImpute or simple median based interpolationwill fail miserably when the data contains extensive amounts of missing entries as in case of GTEx. 

```{r}
robocov_all_genes = get(load("/Users/kushaldey/Documents/Robocov-pages/data/Robocov_Box_all_genes.rda"))
```


```{r fig.height=10, fig.width=10}
corrplot(robocov_all_genes[,,"ENSG00000186517"],   diag = TRUE,
         col = colorRampPalette(c("lightblue4", "lightblue2", "white", "indianred1", "indianred3"))(200),
         tl.pos = "ld", tl.cex = 0.8, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "lower", tl.srt=45)
```

```{r}
plot(eigen(as.matrix(robocov_all_genes[,,"ENSG00000186517"]))$values,
     xlab = "Features (tissues)", ylab = "Eigenvalues",
     main = "Eigenvalue trend for ARHGAP30 gene",
     col = "red",
     pch=20)
```

## Column median imputation 

```{r}
dat = get(load("/Users/kushaldey/Documents/Robocov-pages/data/person_tissue_genes_voom.rda"))
```

```{r}
gene_dat = dat[,,"ENSG00000166819"]
gene_dat_imputed = apply(gene_dat, 2, function(x) {
  y = x
  y[which(is.na(x))] = median(x[!is.na(x)])
  return(y)
})
```

```{r fig.height=10, fig.width=10}
cormat = cor(gene_dat_imputed)
corrplot(cormat,   diag = TRUE,
         col = colorRampPalette(c("lightblue4", "lightblue2", "white", "indianred1", "indianred3"))(200),
         tl.pos = "ld", tl.cex = 0.8, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "lower", tl.srt=45)
```

## SoftImpute imputation

```{r}
impute_method <- "svd"
matc=biScale(dat[,,"ENSG00000166819"],col.scale=TRUE,row.scale=FALSE,trace=TRUE)
fits3=softImpute(matc,rank.max=50,lambda=1,type=impute_method)
fitted_mat <- complete(dat[,,"ENSG00000166819"],fits3,unscale=T)
cormat <- cor(fitted_mat)
```


```{r fig.height=10, fig.width=10}
corrplot(cormat,   diag = TRUE,
         col = colorRampPalette(c("lightblue4", "lightblue2", "white", "indianred1", "indianred3"))(200),
         tl.pos = "ld", tl.cex = 0.8, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "lower", tl.srt=45)
```

```{r}
impute_method <- "svd"
matc=biScale(dat[,,"ENSG00000166819"],col.scale=TRUE,row.scale=FALSE,trace=TRUE)
fits3=softImpute(matc,rank.max=50,lambda=1,type=impute_method)
fitted_mat <- complete(dat[,,"ENSG00000166819"],fits3,unscale=F)
cormat <- cor(fitted_mat)
```

```{r fig.height=10, fig.width=10}
corrplot(cormat,   diag = TRUE,
         col = colorRampPalette(c("lightblue4", "lightblue2", "white", "indianred1", "indianred3"))(200),
         tl.pos = "ld", tl.cex = 0.8, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "lower", tl.srt=45)
```

Pairwise sample correlation matrix 

```{r fig.height=10, fig.width=10}
cormat = cor(gene_dat, use = "pairwise.complete.obs")
corrplot(cormat,   diag = TRUE,
         col = colorRampPalette(c("lightblue4", "lightblue2", "white", "indianred1", "indianred3"))(200),
         tl.pos = "ld", tl.cex = 0.8, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "lower", tl.srt=45)
```

