---
title: "Robocov Blood enrichment compare"
author: "Kushal K. Dey and Alkes Price"
date: "11/23/2019"
output: html_document
---

```{r}
library(readxl)
```

We compare Robospan-All and Robospan-Blood  with other gene scores specific to blood or blood related traits.

```{r}
robospan_all = read.table("/Users/kushaldey/Documents/Robocov-pages/data/Gene_Scores/Robospan_mean.txt")
seg_blood = read.table("/Users/kushaldey/Documents/Robocov-pages/data/Gene_Scores/Fin_GTEx_WholeBlood.txt")
robospan_blood = read.table("/Users/kushaldey/Documents/Robocov-pages/data/Gene_Scores/Robospan_Mean_Blood.txt")
gene_names_gtex = as.character(read.table("/Users/kushaldey/Documents/Robocov-pages/data/Gene_Scores/ensembl_and_hgnc_symbols.txt")[,2])

```

The overlap between the two scores: SEG-GTEx-Blood and Robospan-All is quite weak. Evidence below

```{r}
temp = intersect(robospan_all[,1], seg_blood[,1])
length(temp)/length(robospan_all[,1])
length(temp)/length(seg_blood[,1])
```

Slightly higher overlap of SEG-GTEx-Blood with Robospan-Blood but not super-high enrichment.

```{r}
temp = intersect(robospan_blood[,1], seg_blood[,1])
length(temp)/length(robospan_blood[,1])
length(temp)/length(seg_blood[,1])
```

We remove SEG-GTEx-Blood genes from Robospan-All genes.

```{r}
temp = setdiff(robospan_all[,1], seg_blood[,1])
write.table(temp, file = "/Users/kushaldey/Documents/Robocov-pages/data/Gene_Scores/Robospan_all_comple_SEG_blood.txt",
            row.names = F, col.names = F, quote = F, sep = "\t")
```

## Robopan-All on STRING PPI

```{r}
library(STRINGdb)
string_db <- STRINGdb$new(version="10", species=9606,
                          score_threshold=400, input_directory="")
```

```{r fig.width = 10, fig.height = 10}
pi_drug_targets = data.frame(read_excel("/Users/kushaldey/Documents/Mouse_Humans/data/Pi_approved_drug_targets.xlsx"))
test_genes = unique(pi_drug_targets[-1,4])

df = data.frame(test_genes)
ll <- string_db$map(df, "test_genes", removeUnmappedRows = TRUE )

colors = rep("#FFFFFF", nrow(ll))
colors[match(intersect(robospan_all[,1], ll$test_genes), ll$test_genes)] = "#000000"

colors = as.factor(colors)

payload_id <- string_db$post_payload( ll$STRING_id,
                                      colors)

string_db$plot_network(ll$STRING_id, payload_id=payload_id, required_score = 400)
```


```{r fig.width = 10, fig.height = 10}
pi_drug_targets = data.frame(read_excel("/Users/kushaldey/Documents/Mouse_Humans/data/Pi_approved_drug_targets.xlsx"))
test_genes = unique(pi_drug_targets[-1,4])

df = data.frame(test_genes)
ll <- string_db$map(df, "test_genes", removeUnmappedRows = TRUE )

colors = rep("#FFFFFF", nrow(ll))
colors[match(intersect(robospan_blood[,1], ll$test_genes), ll$test_genes)] = "#000000"

colors = as.factor(colors)

payload_id <- string_db$post_payload( ll$STRING_id,
                                      colors)

string_db$plot_network(ll$STRING_id, payload_id=payload_id, required_score = 400)
```

```{r}
length(intersect(ll$test_genes, robospan_all[,1]))/length(intersect(ll$test_genes, gene_names_gtex))/(1600/14000)
length(intersect(ll$test_genes, robospan_blood[,1]))/length(intersect(ll$test_genes, gene_names_gtex))/(1600/14000)
```

The genes from Robospan-all and Robospan-Blood are not enriched in the approved drug targets. We also compare with PPI-Enhancer and Master regulator genes 

```{r}
master_reg = read.table("/Users/kushaldey/Documents/Mouse_Humans/data/Gene_Scores/master_regulator_genes_Nov11.txt")
ppi_enhancer = read.table("/Users/kushaldey/Documents/Mouse_Humans/data/RWR_Gene_Scores/RWR_SEG_ABC_HOMOD_EDS.txt")
dim(master_reg)
dim(ppi_enhancer)
```

```{r}
length(intersect(master_reg[,1], robospan_all[,1]))/length(intersect(master_reg[,1], gene_names_gtex))/ (1600/13500)
length(intersect(ppi_enhancer[,1], robospan_all[,1]))/length(intersect(ppi_enhancer[,1], gene_names_gtex))/ (1600/13500)
```

Agsin we did not see very high enrichment of these Robospan-Blood and Robospan-All genes with either PPI-Enhancer or Master-Regulator genes. 

This suggests that Robospan scoring scheme is capturingn biologically relevant gene sets orthogonal to either Enhancer-regulated genes or Master-regulator genes. 

## Enriched pathways

The enriched pathways for Robospan-All genes are 

- HSP90 chaperone cycle for steroid hormone receptors (SHR)
- Interferon signaling
- HSF1 activation
- cytokine signaling
- Immune system
- Antigen presentation
- Apoptosis
- Circadian clock

The enriched pathways for Robospan-Blood genes are 

- Immune system
- Innate immune system 
- Neutrophil degranulation 
- Antigen processing and presentation
- Interferon signaling
- HSF1 activation

## Enrichment in housekeeping genes

We next load the list of housekeeping genes.

```{r}
housekeep = read.table("/Users/kushaldey/Documents/Robocov-pages/data/housekeeping_genes.txt")
```

```{r}
length(intersect(housekeep[,1], robospan_all[,1]))/length(intersect(housekeep[,1], gene_names_gtex))/ (1600/14000)
length(intersect(housekeep[,1], robospan_blood[,1]))/length(intersect(housekeep[,1], gene_names_gtex))/ (1600/14000)
```

Again only moderate enrichment of signal among the house-keeping genes.

Next we evaluate constraints on Robospan-All and Robospan-Blood genes.

```{r}
pLI_df = read.table("/Users/kushaldey/Documents/Mouse_Humans/data/Gene_Scores/pLI_genes.txt")
boxplot(pLI_df[match(intersect(robospan_all[,1], pLI_df[,1]), pLI_df[,1]), 2],
pLI_df[match(intersect(robospan_blood[,1], pLI_df[,1]), pLI_df[,1]), 2],
pLI_df[,2])

```

```{r}
pLI_high = pLI_df[which(pLI_df[,2] > 0.99), 1]
length(intersect(pLI_high, robospan_all[,1]))/length(intersect(pLI_high, gene_names_gtex))/ (1600/14000)
length(intersect(pLI_high, robospan_blood[,1]))/length(intersect(pLI_high, gene_names_gtex))/ (1600/14000)
```

```{r}
pLI_low = pLI_df[which(pLI_df[,2] < 1e-08), 1]
length(intersect(pLI_low, robospan_all[,1]))/length(intersect(pLI_low, gene_names_gtex))/ (1600/14000)
length(intersect(pLI_low, robospan_blood[,1]))/length(intersect(pLI_low, gene_names_gtex))/ (1600/14000)
```

```{r}
TF_genes = read.table("/Users/kushaldey/Documents/Mouse_Humans/data/Gene_Scores/TF_genes_curated.txt")
length(intersect(TF_genes[,1], robospan_all[,1]))/length(TF_genes[,1])/ (1600/14000)
length(intersect(TF_genes[,1], robospan_blood[,1]))/length(TF_genes[,1])/(1600/14000)
```

So, far we have not seen that much of a high enrichment of the known gene sets with the Robospan scores; which is interesting. So what is this unique feature that is captured by these scores?

